---
title: "Beeswarm Plots"
format: html
editor: visual
---

## Beeswarm plot

This code was adapted from Stephanie Evergreen's Code for Beeswarm Plots. [This article](https://stephanieevergreen.com/ways-to-make-more-inclusive-data-viz/) goes over the motivation for using a beeswarm plot in detail. The gist is that these are great charts when you want to compare a continuous variable across multiple categories.

## Example code

```{r setup}

library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)

# load data ----
data(iris)
```

Make a vertical beeswarm plot:

```{r}
# vertical beeswarm ----
## make plot ----
plot_vert <- iris %>% 
  ggplot(aes(x = Species, 
             y = Sepal.Length)) +
  geom_point(
    position = position_jitter(w = 1/6, h = 0),
    size = 4, 
    color = "#006cb6", 
    alpha = 2/3
  ) +
  scale_y_continuous(
    limits = c(4, 8),
    breaks = seq(4, 8, 2)
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.25, color = "gray"),
    axis.title = element_blank(),
    axis.text = element_text(face = "bold"),
    axis.text.x = element_text(size = 12)
  ) 
plot_vert
```

Add a title and some formatting to the vertical plot

```{r}
plot_title <- textGrob(
  label = "Sepal Length of Iris Species",
  x = unit(0.2, "lines"), 
  y = unit(0.1, "lines"), 
  hjust = 0, 
  vjust = 0, 
  gp = gpar(fontsize = 18)
)

plot_complete <- arrangeGrob(plot_vert, top = plot_title)
grid.arrange(plot_complete)
```

Make a horizontal beeswarm plot *Note: this is basically the same code as the vertical chart but the x and y variables are switched.*

```{r}
plot_horiz <- iris %>% 
  ggplot(aes(x = Sepal.Length, 
             y = Species)) +
  geom_point(
    position = position_jitter(h = 1/6, w = 0),
    size = 4, 
    color = "#006cb6", 
    alpha = 2/3
  ) +
  scale_x_continuous(
    limits = c(4, 8),
    breaks = seq(4, 8, 2)
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(size = 0.25, color = "gray"),
    axis.title = element_blank(),
    axis.text = element_text(face = "bold"),
    axis.text.y = element_text(size = 12)
  ) + 
  labs(title = "Sepal Length of Iris Species")
plot_horiz
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
